Compiling avoidable deaths codes (as coded by NRS)

### 2021-09-17 compile_avoidable_death_codes.Rmd
Completed compiling avoidable death codes; also added a table showing how code ranges were parsed

# TODO: add code to check length so we can handle decimal codes also!
parse_icd_10_range <- function(code_range) {
  ## helper function for generating code ranges - based on the icd package, function .icd10cm_get_majors_possible()
  if (is.na(code_range) | !str_detect(string = code_range, pattern="\\-")) return(code_range)

@@

```{r}
# TODO: show how the ranges have been parsed just to be sure
avoidable_deaths <-
  avoidable_deaths_raw %>%
  drop_na(condition_group_and_cause) %>%  # remove blank rows used for blank space in the excel file
  separate_rows(icd_10_codes, sep=",\\s*") %>%  # separate comma-separated values into rows so we have one entry per row
  extract(col = icd_10_codes, into = "exclusions", regex = "\\(excl\\.\\s*(.+)\\)", remove = FALSE) %>%  # extract the exclusions and put them in a separate column
  mutate(
    icd_10_codes = str_remove_all(string = icd_10_codes, pattern = "\\s*\\(excl.+\\)"),  # remove the exclusions from the code ranges
    icd_10_codes = map_chr(.x = icd_10_codes, .f = ~parse_icd_10_range(.x))
  )
```

@@

>>>>

parse_icd_10_range <- function(code_range) {
  ## helper function for generating code ranges - based on the icd package, function .icd10cm_get_majors_possible()
  if (is.na(code_range) | !str_detect(string = code_range, pattern="\\-")) return(code_range)

@@

```{r}
avoidable_deaths <-
  avoidable_deaths_raw %>%
  drop_na(condition_group_and_cause) %>%  # remove blank rows used for blank space in the excel file
  mutate(
    group = if_else(condition = is.na(icd_10_codes), true = condition_group_and_cause, false = NA_character_),
    original_codes = icd_10_codes
    ) %>%
  fill(group, .direction = "down") %>%  # forward-fill group
  drop_na(icd_10_codes) %>%
  separate_rows(icd_10_codes, sep=",\\s*", ) %>%  # separate comma-separated values into rows so we have one entry per row
  extract(col = icd_10_codes, into = "exclusions", regex = "\\(excl\\.\\s*(.+)\\)", remove = FALSE) %>%  # extract the exclusions and put them in a separate column
  mutate(
    icd_10_codes = str_remove_all(string = icd_10_codes, pattern = "\\s*\\(excl.+\\)"),  # remove the exclusions from the code ranges
    icd_10_codes = map_chr(.x = icd_10_codes, .f = ~parse_icd_10_range(.x))
  ) %>%
  separate_rows(icd_10_codes, sep=",\\s*") %>%  # separate comma-separated values again now that we've converted them from the ranges
  separate(col = age, into = c("age_min", "age_max") ,sep = "\\-", remove = TRUE) %>%
  mutate(
    treatable = case_when(  # make more explicit the treatable/preventable distinction
      str_detect(treatable, pattern = "50") ~ 0.5,
      str_detect(treatable, pattern = "\\*") ~ 1,
      TRUE ~ 0
    ),
    preventable = case_when(
      str_detect(preventable, pattern = "50") ~ 0.5,
      str_detect(preventable, pattern = "\\•") ~ 1,
      TRUE ~ 0
    ),
  ) %>%
  relocate(group, condition_group_and_cause, original_codes)
```

## Checking parsing of ranges

```{r}
avoidable_deaths_raw %>%
  drop_na(condition_group_and_cause) %>%  # remove blank rows used for blank space in the excel file
  separate_rows(icd_10_codes, sep=",\\s*") %>%  # separate comma-separated values into rows so we have one entry per row
  extract(col = icd_10_codes, into = "exclusions", regex = "\\(excl\\.\\s*(.+)\\)", remove = FALSE) %>%  # extract the exclusions and put them in a separate column
  filter(str_detect(string = icd_10_codes, pattern = "\\-")) %>%
  mutate(
    icd_10_codes = str_remove_all(string = icd_10_codes, pattern = "\\s*\\(excl.+\\)"),  # remove the exclusions from the code ranges
    icd_10_codes_parsed = map_chr(.x = icd_10_codes, .f = ~parse_icd_10_range(.x))
  ) %>%
  select(icd_10_codes, icd_10_codes_parsed) %>% 
  kable(caption = "Parsing of codes defined as ranges.")
```

@@

# Save as `.csv`

```{r}
if (!dir.exists("./avoidable_deaths")) dir.create("./avoidable_deaths")
write_csv(avoidable_deaths, file = "./avoidable_deaths/avoidable_deaths_2019_inclusions.csv")
```
